---
title: "OECD Country comparisons"
date: "`r Sys.Date()`"
output:
  rmdformats::html_clean:
    highlight: kate
---

# Neve`r`ending race 

When looking at OECD economic indicators, what comes to mind is rankings between countries.
What follows is an attempt at visualizing this ...neverending race.


## The `R` libraries

Which are the libraries we will need and why

`# devtools::install_github("jbkunst/highcharter")`   
`# devtools::install_github("ramnathv/rblocks")`    
`# devtools::install_github("woobe/rPlotter")`    

```{r include=T, warning=F, echo=T, message=F}
library(gdata) # reading xls
library(dplyr) # data manipulation
library(highcharter) # visualization
library(stringdist) # some fuzzy text matching
library(rvest) # getting data from the web
library(rPlotter) # whats the color in the image?
library(knitr) # kniting to pass the time
library(magrittr) # human readable code

```

## User input

The OECD indicators we will look at, are `Productivity` and `Unemployment`, for 10 counties in Europe, starting with productivity.

```{r include=T, warning=F, echo=T, message=F}
indicator <- c("GDP per hour worked"=".GDPHRWKD.TOT.IDX2010.A", "Unemployment"=".UNEMP.TOT.PC_LF.A")

### End user selection ###
ind <- indicator[1]
YEARS <- 2001:2015
selected_countries <- c("Austria", "Belgium", "Germany", "Spain", "Finland", 
                        "France", "United Kingdom","Greece", "Ireland", "Iceland")

```

You can find a list of 234 countries to choose from within this linked [excel file](http://www.oecd.org/migration/mig/34107835.xls).

# The data

OECD uses abbreviations for the countries, lets use the mapping that is out there

```{r include=T, warning=F, echo=T, message=F}

# Get the countries and country abbreviations from OECD
cntry_Abb <- read.xls("http://www.oecd.org/migration/mig/34107835.xls")[, 1:2]
cntry_Abb$CODE <- gsub(".*-", "", cntry_Abb$CODE)
cntry_Abb$Country <- as.character(cntry_Abb$Country)


# Fixing some mistakes
cntry_Abb$Country[cntry_Abb$CODE=="SO"] <- "South Korea"
cntry_Abb$Country[cntry_Abb$CODE=="NS"] <- "North and South Korea"
cntry_Abb <- rbind(cntry_Abb, cbind(CODE="KOR", Country="Korea"))
cntry_Abb$Country[cntry_Abb$CODE=="UGA"] <- "Uganda"
cntry_Abb$Country[cntry_Abb$CODE=="VAT"] <- "Vatican"


# Get the selected indicator figures for the selected countries
indic_url <- "https://stats.oecd.org/sdmx-json/data/DP_LIVE/%s%s/OECD?contentType=csv&detail=code&separator=comma&csv-lang=en&startPeriod=1998&endPeriod=2016"
in_url <- sprintf(indic_url, paste0(cntry_Abb$CODE[tolower(cntry_Abb$Country) %in% tolower(selected_countries)], collapse="+"), ind)
df <- read.csv(in_url)

```

Some data manipulation

```{r include=T, warning=F, echo=T, message=F}
names(df) <- c("CODE", "INDICATOR","SUBJECT","MEASURE","FREQUENCY","Year","Value","Flag.Codes")

df %<>% left_join(cntry_Abb) %>% 
  select(Country, Year, Value) %>% mutate(Value=round(Value, 1)) %>% 
  filter(Country %in% selected_countries) %>% arrange(desc(Value)) %>% 
  group_by(Year) %>% mutate(Rank=1, Rank=as.integer(cumsum(Rank))) %>% 
  ungroup() %>% {.[complete.cases(.),]}


kable(head(df,15), digits = 2)
```



<br>
Lets fetch some flags, and fuzzy match them to the countries

```{r include=T, warning=F, echo=T, message=F}

# Getting flag icons
flagUrl <- "http://www.iconarchive.com/show/round-world-flags-icons-by-custom-icon-design.%d.html"
h <- lapply(sprintf(flagUrl, 1:6), function(x) read_html(x) %>% html_nodes(".icondetail img") %>% 
              xml_attr("src"))
flagUrls <- unlist(h)
flagCountries <- sub(".*/(.*)\\.png", "\\1", flagUrls) %>% 
  gsub("-icon", "", ., fixed=TRUE) %>% gsub("-", " ", .)

# Some fuzzy matching on the countries
df$urls <- paste0("url(",flagUrls[amatch(df$Country, flagCountries,  method = "cosine")], ")")

# extract a color from the flag
df$flagcolor <- gsub("url\\(|\\)", "", df$urls)
flagcolor <- sapply(unique(df$flagcolor), function(x) extract_colours(x, 3)[2])
df$flagcolor <- flagcolor[df$flagcolor]

```


# Prep the data

```{r include=T, warning=F, echo=T, message=F}

# prepare data for the chart
df %<>% filter(Year %in% YEARS) %>% select(-Value)
df$z <- 1


cntr <- unique(df$Country)
df1 <- lapply(cntr, function(i){
  
  df1 <- df %>% filter(Country == i) %>% arrange(Year) %>% 
  {do.call(rbind, replicate(dim(.)[1], ., simplify = FALSE))} %>% 
    group_by(Country, Year) %>% mutate(key=1, key=cumsum(key)) %>%
    group_by(key) %>% mutate(key2=1, key2=cumsum(key2)) %>% 
    ungroup() %>% mutate(Rank=ifelse(key<key2, NA, Rank))
  
  df1 %>% filter(key==1) %>% select(-key) %>% 
    left_join(df1 %>% group_by(key2) %>% 
                do(sequence = list_parse(select(.,x=Year, y=Rank, z=z))) %>% 
                ungroup(), by="key2") %>% select(-key2)
})


bubble_df <- lapply(cntr, function(i){
  
  left_join(df %>% filter(Country==i) %>% arrange(Year) %>% slice(1), 
            df %>% filter(Country==i) %>% arrange(Year) %>% group_by(Country) %>% 
              do(sequence = list_parse(select(., x=Year, y=Rank, z=z))))

})

```



# Let the race begin


## Productivity

``` {r include=T, warning=F, echo=T, message=F}
# building the highchart
hc_exp <- paste("highchart() %>% ", 
                paste("hc_add_series(data = df1[[", 1:length(cntr), "]], 
                      type = \"line\", hcaes(x = Year, y = Rank), 
                      dashStyle='solid', enableMouseTracking=FALSE, 
                      color=df1[[", 1:length(cntr), "]]$flagcolor[1],
                      lineWidth=10,  marker=list(radius=4, fillColor=\"white\"), 
                      name=\" \")", collapse=" %>% "), " %>% ",
                
                paste("hc_add_series(bubble_df[[", 1:length(cntr), "]], 
                      type = \"bubble\", hcaes(x = Year, y = Rank, size = z), 
                      name=bubble_df[[", 1:length(cntr), "]]$Country, 
                      marker=list(symbol=bubble_df[[", 1:length(cntr), "]]$urls, 
                      radius=2))", collapse=" %>% "), " %>% ", 
                
                "hc_yAxis(min = 0, max = ", length(cntr)+1,", reversed = TRUE, tickInterval=1, labels=\"\") %>%
                hc_xAxis(min = 2001, max = 2016, tickInterval=1) %>%", 
                
                " hc_plotOptions(bubble=list(maxSize='10%')) %>% ",
                
                "hc_motion(enabled = TRUE, series = c(0:", (2*length(cntr)-1),"), startIndex = 0, updateInterval = 1, labels=\"<br><br><br><br><br>\") %>% 
                  hc_chart(backgroundColor = '#eaecef')"
                
                )

# What is that:
# cat(hc_exp)

```

Higher rank higher productivity

`eval` to the the rescue
```{r include=T, warning=F, echo=T, message=F}
eval(parse(text=hc_exp))
```
<br>
As we can see because of the Subprime mortgage crisis there was a lot of movement in the years that followed 2008.
Greece is another story by itself.

## Unemployment

higher rank, higher unemployment
```{r include=F, warning=F, echo=T, message=F}
ind <- indicator[2]
YEARS <- 2003:2015

indic_url <- "https://stats.oecd.org/sdmx-json/data/DP_LIVE/%s%s/OECD?contentType=csv&detail=code&separator=comma&csv-lang=en&startPeriod=1998&endPeriod=2016"
in_url <- sprintf(indic_url, paste0(cntry_Abb$CODE[tolower(cntry_Abb$Country) %in% tolower(selected_countries)], collapse="+"), ind)
df <- read.csv(in_url)

names(df) <- c("CODE", "INDICATOR","SUBJECT","MEASURE","FREQUENCY","Year","Value","Flag.Codes")

df %<>% left_join(cntry_Abb) %>% 
  select(Country, Year, Value) %>% mutate(Value=round(Value, 1)) %>% 
  filter(Country %in% selected_countries) %>% arrange(desc(Value)) %>% 
  group_by(Year) %>% mutate(Rank=1, Rank=as.integer(cumsum(Rank))) %>% 
  ungroup() %>% {.[complete.cases(.),]}


# Getting flag icons
flagUrl <- "http://www.iconarchive.com/show/round-world-flags-icons-by-custom-icon-design.%d.html"
h <- lapply(sprintf(flagUrl, 1:6), function(x) read_html(x) %>% html_nodes(".icondetail img") %>% xml_attr("src"))
flagUrls <- unlist(h)
flagCountries <- sub(".*/(.*)\\.png", "\\1", flagUrls) %>% gsub("-icon", "", ., fixed=TRUE) %>% gsub("-", " ", .)

# Some fuzzy matching on the countries
df$urls <- paste0("url(",flagUrls[amatch(df$Country, flagCountries,  method = "cosine")], ")")

# extract a color from the flag
df$flagcolor <- gsub("url\\(|\\)", "", df$urls)
flagcolor <- sapply(unique(df$flagcolor), function(x) extract_colours(x, 3)[2])
df$flagcolor <- flagcolor[df$flagcolor]

# prepare data for the chart
df %<>% filter(Year %in% YEARS) %>% select(-Value)
df$z <- 1


cntr <- unique(df$Country)
df1 <- lapply(cntr, function(i){
  
  df1 <- df %>% filter(Country == i) %>% arrange(Year) %>% 
  {do.call(rbind, replicate(dim(.)[1], ., simplify = FALSE))} %>% 
    group_by(Country, Year) %>% mutate(key=1, key=cumsum(key)) %>%
    group_by(key) %>% mutate(key2=1, key2=cumsum(key2)) %>% 
    ungroup() %>% mutate(Rank=ifelse(key<key2, NA, Rank))
  
  df1 %>% filter(key==1) %>% select(-key) %>% 
    left_join(df1 %>% group_by(key2) %>% 
                do(sequence = list_parse(select(.,x=Year, y=Rank, z=z))) %>% 
                ungroup(), by="key2") %>% select(-key2)
})


bubble_df <- lapply(cntr, function(i){
  
  left_join(df %>% filter(Country==i) %>% arrange(Year) %>% slice(1), 
            df %>% filter(Country==i) %>% arrange(Year) %>% group_by(Country) %>% 
              do(sequence = list_parse(select(., x=Year, y=Rank, z=z))))

})

hc_unem <- paste("highchart() %>% 
                 ", 
                paste("hc_add_series(data = df1[[", 1:length(cntr), "]], type = \"line\", hcaes(x = Year, y = Rank), 
                      dashStyle='solid', enableMouseTracking=FALSE, color=df1[[", 1:length(cntr), "]]$flagcolor[1],
                      lineWidth=10,  marker=list(radius=4, fillColor=\"white\"), name=\" \")", collapse=" %>% 
                      "), " %>% 
                ",
                
                paste("hc_add_series(bubble_df[[", 1:length(cntr), "]], type = \"bubble\", hcaes(x = Year, y = Rank, size = z), 
                      name=bubble_df[[", 1:length(cntr), "]]$Country, 
                      marker=list(symbol=bubble_df[[", 1:length(cntr), "]]$urls, radius=2))", collapse=" %>% 
                      "), " %>% 
                ", 
                
                "hc_yAxis(min = 0, max = ", length(cntr)+1,", reversed = TRUE, tickInterval=1, labels=\"\") %>%

                hc_xAxis(min = 2003, max = 2016, tickInterval=1) %>%", 
                
                " hc_plotOptions(bubble=list(maxSize='10%')) %>% ",
                
                "hc_motion(enabled = TRUE, series = c(0:", (2*length(cntr)-1),"), startIndex = 0, updateInterval = 1, labels = 1:15) %>% 
                  hc_chart(backgroundColor = '#eaecef')"
                
                )

```


```{r}
eval(parse(text=hc_unem))
```
